### 2020-05-22 paper/appendix.tex
shorthand

\documentclass[format=acmsmall, review=true]{acmart}
\usepackage{acm-tweaks}
\usepackage{booktabs} % For formal tables
\usepackage[ruled]{algorithm2e} % For algorithms
\usepackage{subcaption}
\newcommand{\xhdr}[1]{\vspace{1mm} \noindent{\bf #1}}

\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}
% Choose a citation style by commenting/uncommenting the appropriate line:
%\setcitestyle{authoryear}
\setcitestyle{acmnumeric}

\newtheorem{finding}{Finding}

% Title. Note the optional short title for running heads. In the interest of anonymization, please do not include any acknowledgements.
\newcommand\PaperTitle{Appendix For Deconstructing the Filter Bubble: Consumer Decision-Making and Recommender Systems}
\title[\PaperTitle]{\PaperTitle}

\begin{document}

\maketitle
\addtocounter{figure}{+4}
\section{Additional Figures}

This section contains figures that were omitted from the main text due to space constraints. Thus, the figures in this section are computed utilizing the same parameter values described in the main text for $N = 200$.

\begin{figure}[ht]
\caption{Relationship between $\rho$ and User Welfare, Diversity, $N = 200$}\label{fig:diversity_welfare_correlation}
\begin{subfigure}{0.3\linewidth}
  \includegraphics[width=1.0\linewidth]{figures/rho_diversity_N_200_T_20.pdf}
\end{subfigure}
\ContinuedFloat
\begin{subfigure}{0.3\linewidth}
  \includegraphics[width=1.0\linewidth]{figures/rho_welfare_N_200_T_20.pdf}
\end{subfigure}
\caption*{\scriptsize Notes: The figure on the left displays the relationship between $\rho$, the correlation between valuations of items, and overall consumed product diversity. The figure on the right displays the relationship between $\rho$ and overall welfare.}
\end{figure}

\addtocounter{figure}{-1}

\begin{figure}[H]
\caption{Relationship between Local Consumption and $\rho$, $N = 200$}
\begin{subfigure}{.3\textwidth}
\includegraphics[width=\linewidth]{figures/rho_consumption_dist_N_200T_20.pdf}
\end{subfigure}
\begin{subfigure}{.3\textwidth}
\includegraphics[width=\linewidth]{figures/rho_consumption_dist_N_200T_20_partial.pdf}
\end{subfigure}\\
\begin{subfigure}{.3\textwidth}
\includegraphics[width=\linewidth]{figures/rho_consumption_dist_N_200T_20_omni.pdf}\\
\end{subfigure}
\caption*{\scriptsize Notes: Each figure plots the average consecutive consumption distance across time as the inherent correlation between the valuation of the items, $\rho$. The top left displays the no recommendation regime, the top right displays the recommendation regime, and the bottom displays the oracle regime.}
\label{fig:local_consumption_across_rho}
\end{figure}
\addtocounter{figure}{-1}

\begin{figure}[H]
\caption{Relationship between Local Consumption and $\gamma$, $N = 200$}
\begin{subfigure}{.3\textwidth}
\includegraphics[width=\linewidth]{figures/gamma_consumption_dist_N_200T_20.pdf}
\end{subfigure}
\begin{subfigure}{.3\textwidth}
\includegraphics[width=\linewidth]{figures/gamma_consumption_dist_N_200T_20_partial.pdf}
\end{subfigure}
\begin{subfigure}{.3\textwidth}
\includegraphics[width=\linewidth]{figures/gamma_consumption_dist_N_200T_20_omni.pdf}\\
\end{subfigure}%
\caption*{\scriptsize Notes: Each figure plots the average consecutive consumption distance across time as the risk aversion level of users, $\gamma$, varies. The top left displays the no recommendation regime, the top right displays the recommendation regime, and the bottom displays the oracle regime.}
\label{fig:no_rec_risk_aversion}
\end{figure}
\addtocounter{figure}{-1}

\begin{figure}[ht]
\caption{Relationship between $\gamma$ and Diversity, $N = 200$}
\includegraphics[width=.45\linewidth]{figures/gamma_diversity_N_200_T_20}
\caption*{\scriptsize Notes: This figure displays the value of the diversity measure as we vary the risk aversion level, $\gamma$.}\label{fig:cor_homo}
\end{figure}

\begin{figure}[ht]
\caption{Relationship between $\beta$ and User Welfare, Diversity, $N = 200$}\label{fig:diversity_welfare_correlation}
\begin{subfigure}{.45\linewidth}
  \includegraphics[width=.8\linewidth]{figures/beta_diversity_N_200_T_20.pdf}
\end{subfigure}
\ContinuedFloat
\begin{subfigure}{.45\linewidth}
  \includegraphics[width=.8\linewidth]{figures/beta_welfare_N_200_T_20.pdf}
\end{subfigure}
\caption*{\scriptsize Notes: The figure on the left displays the relationship between $\beta$, the strength of the common value component, and overall consumed product diversity. The figure on the right displays the relationship between $\beta$ and overall welfare.}
\end{figure}

\addtocounter{figure}{-1}

% Bibliography

\end{document}